\makeatletter
\def\input@path{{../.format/}}
\makeatother

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\input{macros}

\begin{document}

\section{Prelude to Metalogical Inference}

The "inexplicable yet self-evident" nature of consciousness finds its most precise articulation in the language of formal logic. Among mathematical disciplines, logic uniquely addresses both derivation and foundation: the rules for proving statements and the axioms accepted without proof. For comparison, other sciences rely on various mathematical tools—physics employs calculus, differential geometry, and tensor analysis \cite{Landau1976,Arnold1989,Misner1973}; neuroscience utilizes dynamical systems, information theory, and statistical mechanics \cite{Dayan2001,Friston2010,Hopfield1982}; computer science leverages automata theory and complexity theory \cite{Sipser2012}. But only logic provides the formal structure to express what cannot be derived yet must be accepted as true.

The formal mapping is precise: the "self-evident" aspect of conscious experience corresponds to axiomaticity, statements accepted as foundational without proof; the "inexplicable" aspect corresponds to undecidability, propositions neither provable nor disprovable within a given formal system \cite{Mendelson2009}. Gödel's incompleteness theorems established that in any sufficiently powerful formal system containing basic arithmetic, there exist true statements unprovable within that system \cite{Godel1931}. This limitation wasn't a failure of Hilbert's program to establish mathematics on unshakable foundations \cite{Hilbert1922}, but a necessary feature of expressive formal systems. Consciousness exhibits this same characteristic: its existence is certain yet formally inexplicable.

We now construct a counterfactual narrative beginning with Gödel's work. After establishing his revolutionary results through arithmetization, the provability predicate, and diagonalization, historical Gödel demonstrated the limitations of formal systems. But what if he had instead explored how formal systems might evolve when confronting their own limitations? This alternative intellectual path—unexplored in our history but rich with possibility—leads us to the concept of Metalogical Inference.

\setcounter{definition}{1}
\begin{cdefinition}
    \textbf{Metalogical Inference} is the process by which a logical system recognizes truths that transcend its current formal framework and adjusts its structure to accommodate them while preserving the semantics of logic on which it is founded.
\end{cdefinition}

%-----------------------------------------------------------------------------------------------------

\begin{thebibliography}{9}

\bibitem{Godel1931}
Gödel, K.: On Formally Undecidable Propositions of *Principia Mathematica* and Related Systems I.  
Originally published as: *Über formal unentscheidbare Sätze der Principia Mathematica und verwandter Systeme I*. \textit{Monatshefte für Mathematik und Physik}, \textbf{38}(1), 173–198 (1931).

\bibitem{Mendelson2009}
Mendelson, E.: \textit{Introduction to Mathematical Logic}. 5th ed. (2009)

\bibitem{Hilbert1922}
Hilbert, D.: Die Grundlagen der Mathematik. (1922)

\bibitem{Misner1973}
Misner, C.W., Thorne, K.S., Wheeler, J.A.: \textit{Gravitation}. W.H. Freeman (1973)

\bibitem{Landau1976}
Landau, L.D., Lifshitz, E.M.: \textit{The Classical Theory of Fields}. (1976)

\bibitem{Arnold1989}
Arnold, V.I.: \textit{Mathematical Methods of Classical Mechanics}. (1989)

\bibitem{Hopfield1982}
Hopfield, J.J.: Neural networks and physical systems... \textit{PNAS} \textbf{79}(8), 2554–2558 (1982)

\bibitem{Sipser2012}
Sipser, M.: \textit{Introduction to the Theory of Computation}. Cengage (2012)

\bibitem{Dayan2001}
Dayan, P., Abbott, L.F.: \textit{Theoretical Neuroscience: Computational and Mathematical Modeling of Neural Systems}. MIT Press (2001)

\bibitem{Friston2010}
Friston, K.: The free-energy principle: a unified brain theory? \textit{Nature Reviews Neuroscience} \textbf{11}(2), 127--138 (2010)

\end{thebibliography}

\end{document}
