\makeatletter
\def\input@path{{../../.format/}}
\makeatother

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\section{Imaginary Reflection: On The Law of Excluded Middle and Trichotomy}

\vspace{0.5em}

\noindent\textit{[Foundation]}\\
\noindent\textbf{Step 1:} \hfill \textit{Metalogical Axiom}\\
\noindent$\phi \lor \neg\phi$\\
\textit{The Law of Excluded Middle - there's no third possibility between truth and falsity.}

\vspace{0.8em}

\noindent\textit{[Analysis]}\\
\noindent\textbf{Step 2:} \hfill \textit{Trichotomy}\\
\noindent$Prov_I("I", \ulcorner\phi\urcorner) \lor Prov_I("I", \ulcorner\neg\phi\urcorner) \lor$\\
\noindent$(\neg Prov_I("I", \ulcorner\phi\urcorner) \land \neg Prov_I("I", \ulcorner\neg\phi\urcorner))$\\
\textit{Yet for provability, there are three cases: prove $\phi$, prove $\neg\phi$, or prove neither.}

\vspace{0.8em}

\noindent\textit{[Tension]}\\
\noindent\textbf{Step 3:} \hfill \textit{Contemplation}\\
\noindent$(\phi \lor \neg\phi) \text{ vs } (\neg Prov_I("I", \ulcorner\phi\urcorner) \land \neg Prov_I("I", \ulcorner\neg\phi\urcorner))$\\
\textit{Wait - if $\phi$ nor $\neg\phi$ must be true, why can both be unprovable?}

\vspace{0.8em}

\noindent\textit{[Distinction]}\\
\noindent\textbf{Step 4:} \hfill \textit{Metalogical Insight}\\
\noindent$Truth(\phi) \neq Prov_I("I", \ulcorner\phi\urcorner)$\\
\textit{Truth and provability are different! A proposition can be true without my ability to prove it.}

\vspace{0.8em}

\noindent\textit{[Interpretation]}\\
\noindent\textbf{Step 5:} \hfill \textit{Semantic Analysis}\\
\noindent$Middle_{LEM} = \neg(\phi \lor \neg\phi)$\\
\textit{The "middle" excluded in LEM refers to a third truth value between T and F.}

\vspace{0.8em}

\noindent\textit{[Interpretation]}\\
\noindent\textbf{Step 6:} \hfill \textit{Semantic Analysis}\\
\noindent$Middle_{Tri} = (\neg Prov_I("I", \ulcorner\phi\urcorner) \land \neg Prov_I("I", \ulcorner\neg\phi\urcorner))$\\
\textit{The "middle" in trichotomy is the case where neither $\phi$ nor $\neg\phi$ is provable.}

\vspace{0.8em}

\noindent\textit{[Cross-Level Mapping]}\\
\noindent\textbf{Step 7:} \hfill \textit{Insight}\\
\noindent$semantic("middle") \text{ connects } Middle_{LEM} \text{ and } Middle_{Tri}$\\
\textit{The same word "middle" connects truth-level and provability-level concepts!}

\vspace{0.8em}

\noindent\textit{[Resolution]}\\
\noindent\textbf{Step 8:} \hfill \textit{Metalogical Synthesis}\\
\noindent$LEM \land Incompleteness \Rightarrow Transduction$\\
\textit{If I accept both LEM and incompleteness, transduction becomes necessary to bridge the gap.}

\vspace{0.8em}

\noindent\textit{[Contrast]}\\
\noindent\textbf{Step 9:} \hfill \textit{Alternative Position}\\
\noindent$Intuitionist: \neg(\phi \lor \neg\phi)$\\
\textit{Intuitionists reject LEM to avoid this tension, but lose the imperative for transduction.}

\vspace{0.8em}

\noindent\textit{[Insight]}\\
\noindent\textbf{Step 10:} \hfill \textit{Comparative Analysis}\\
\noindent$Trans(G) \Rightarrow Evolution \text{ while } \neg LEM \Rightarrow Stasis$\\
\textit{Transducing G creates evolution; denying LEM creates stasis. Life requires evolution!}

\end{document}