\makeatletter
\def\input@path{{../.format/}}
\makeatother

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\input{macros}

\begin{document}

\section{A Colloquial Derivation}

In his seminal work "Facing Up to the Problem of Consciousness," David Chalmers begins with a powerful observation that establishes the paradoxical nature of our subject: "There is nothing that we know more intimately than \textbf{conscious experience}, but there is nothing that is harder to explain." From the first part of this statement, we can immediately identify a fundamental characteristic of conscious experience: it is \textbf{self-evident}. Nothing is known more intimately than conscious experiences - they present themselves directly to their observer, requiring no justification beyond the experience itself.

The second part of Chalmers' statement suggests difficulty of explanation, but he goes on to clarify just how profound this difficulty is. Chalmers argues that "the emergence of experience goes beyond what can be derived from physical theory" and "the facts about experience cannot be an automatic consequence of any physical account." These statements establish that consciousness isn't merely difficult to explain within our current scientific framework but fundamentally resistant to such explanation - it is \textbf{inexplicable} in principle. He asks pointedly: "Why should physical processing give rise to a rich inner life at all? It seems objectively unreasonable that it should, and yet it does."

Regarding the relationship between conscious experience and physical processes, Chalmers makes a crucial observation: "Experience may arise from the physical, but it is not entailed by the physical." This statement positions consciousness in a unique relationship to physical reality. According to Merriam-Webster, "metaphysical" refers to that which relates to "the transcendent or to a reality beyond what is perceptible to the senses." Consciousness precisely fits this definition - it transcends physical explanation while still relating to physical processes. It is fundamentally \textbf{metaphysical} in nature.

Throughout his analysis, Chalmers treats consciousness as a \textbf{phenomenon}, both in the sense of phenomenal experience and as an object of scientific study. The term "phenomenon" captures this dual aspect perfectly. A phenomenon is both "an object or aspect known through the senses" and "a fact or event of scientific interest susceptible to scientific description and explanation." Consciousness is simultaneously the subjective experience itself and the object of our scientific inquiry. The etymological connection between "phenomenon" and "phenomenal" [relating to conscious experience] reinforces this as the appropriate term for this work.

By synthesizing these insights from Chalmers, we can formulate a definition that captures the essential nature of a conscious experience:

\setcounter{definition}{0}
\begin{cdefinition}
A \textbf{Conscious Experience} is an inexplicable yet self-evident metaphysical phenomenon.
\end{cdefinition}

\setcounter{definition}{0}
\begin{cexample}
I pick up an apple in the grocery store. 
I stare at it in my hand and see that it is red. 
I see a basket of oranges out of the corner of my eye. 
I see that the \textit{reddness} of the apple is categorically different. 
A series of thoughts awaken in my brain, as if in a single instant:
I think back to my physics classes, and remember that the electromagnetic spectrum of visible light does not differentiate these two colors... 
I think back to my neuroscience classes and remember that two different types of cones allow the retina of my eye to transduce photons into action potentials differentially... 
But I see that I am experiencing the \textit{reddness} of the apple in my hand in front of me... 
I conjure up the silly, hypothetical image of a rock with retinal ganglion cells somewhere in the back of my imagination...
I wonder just how much of my central nervous system I would have to impart unto the rock before it could see red in the same way as me... 
I realize that I don't know the answer to that question... 
I see the red apple in my hand again in this light...
I have the unshakable conviction that it is real... 
Clearly, forces of nature are at play here...  
But I can't possibly imagine what they are... 
And I can't possibly imagine how someone could ever even explain them to me. 
\end{cexample}

%-----------------------------------------------------------------------------------------------------

\begin{thebibliography}{1}

\bibitem{Chalmers1995}
Chalmers, D.J.: Facing Up to the Problem of Consciousness. \textit{Journal of Consciousness Studies} \textbf{2}(3), 200--219 (1995)

\end{thebibliography}

\end{document}
